---
title: "Loading your SS3 results into FLR with ss3om"
author: Iago Mosqueira (EC JRC) <iago.mosqueira@ec.europa.eu>
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{ss3om}
  %\VignetteEncoding{UTF-8}
tags:
abstract:
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
---

```{r, pkgs, echo=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
  fig.width=5, fig.height=4.5)
```

The *ss3om* package has been developed to bring the output of the Stock Synthesis (**SS**) integrated stock assessment platform [@Methot_2013] into the **FLR** framework [@Kell_2007]. The package initial objective was to allows the use of **SS** model runs to construct Operating Models (OM) for Management Strategy Evaluation (MSE) of candidate Management Procedures (MP). The outputs of model runs are used to construct object of the various classes employed when running evaluation of management procedures using the FLR methods, particularly those defined in the [`mse`](https://flr-project.org/mse) package. Initial and current work on these code is related to the evaluation of management procedures for some stocks managed by the [Indian Ocean Tuna Commission](https://iotc.org/) (IOTC), but has been extended for other stocks and model configurations.

The MPs evaluated using these OMs are generally based on simpler estimators of stock status or trends, so no attempt is made here to provide facilities for running **SS** as part of a management procedure. For simulations involving **SS** as a status indicator, please take a look at the [ss3sim](https://github.com/ss3sim/ss3sim) package.

This package has been used on a number of **SS** models (see below), for both versions 3.24 and 3.30. Stock synthesis is a very flexible framework, able to allow construction of a large number of different models. Not all of them can currently be loaded into **FLR** using *ss3om*. If your model fails to load, or the objects do not quite hold the right information, please do not hesitate contacting us to see if the code could be extended to work on it. However, development is now focused on models coming from **SS** 3.30, including some of the new features it has brought. Nevertheless, interaction with the output and input files is taken care by the [r4ss](https://github.com/r4ss/r4ss) package. So as long as *r4ss* can load your model, there is a good chance we can load it into **FLR**.

# Installation


The package is available from the **FLR** repository, and can be installed, including all its dependencies, by calling

```{r, eval=FALSE}
install.packages("ss3om", repos=structure(c(CRAN="https://cloud.r-project.org/",
  FLR="https://flr-project.org/R")))
```

You can also install the development version directly from the [github repository](https://github.com/flr/ss3om). For this an updated version of [r4ss](https://github.com/r4ss/r4ss) is needed:

```{r, eval=FALSE}
remotes::install_github("r4ss/r4ss")
remotes::install_github("flr/ss3om")
```

# Using `read` functions to load SS3 results

```{r loadpkg}
library(ss3om)
```

```{r setpwd}
dir <- system.file("ext-data/herring", package="ss3om")
```

## *FLStock*

```{r readfls}
her <- readFLSss3(dir, name="HER")
range(her) <- c(minfbar=3, maxfbar=7)

plot(her)
```

| FLR class | From folder | From list   |               |
|-----------|-------------|-------------|---------------|
| FLStock   | readFLSss3  | buildFLSss3 | buildFLSss330 |
| FLSR      | readFLSRss3 | buildFLSss3 | buildFLSss330 |


```
        SS_output         buildFL*
folder -----------> list ---------> FL*
       `----------- readFL -------->|

```

# The `sol` dataset

The package ships with `sol`, a full `r4ss::SS_output` list obtained from a **SS** 3.30 assessment of a sole stock, and stored as a `data(sol)` dataset. The original model files, used by the `read*` functions below, are also distributed with the package under `system.file("ext-data/sol", package="ss3om")`. This dataset is used in this section to demonstrate every `read` and `build` function in *ss3om*.

```{r solpkgs, echo=TRUE}
data(sol)

class(sol)
sol$SS_version
```

## `read` functions on `sol`

The `read*` functions load a **SS** run directly from its output folder, calling `readOutputss3()` internally and then dispatching to the relevant `build*` function depending on the **SS** version of the model.

```{r solread, echo=TRUE}
dir <- system.file("ext-data/sol", package="ss3om")
```

| Function        | Output class | Description                                    |
|------------------|--------------|-------------------------------------------------|
| readOutputss3   | *list*       | Loads Report, CompReport and covar files with `r4ss::SS_output` |
| readFLSss3      | *FLStock*    | Stock and catch time series                     |
| readFLSRss3     | *FLSR*       | Stock-recruitment relationship                  |
| readFLRPss3     | *FLPar*      | Reference points (B0, R0, SBMSY, FMSY, MSY, ...) |
| readKobess3     | *FLQuants*   | B/BMSY and F/FMSY time series (Kobe plot)        |
| readRESss3      | *data.table* | Time series of stock summary results            |
| readFLIBss3     | *FLIndices*  | Abundance indices, as `FLIndexBiomass`           |
| readFLomss3     | *FLom*       | Operating model, stock + stock-recruit + refpts  |
| readOMSss3      | *list*       | List of elements used to build an operating model|
| readCDss3       | *list*       | `ctl` and `dat` input files, via `r4ss`          |

```{r solreadfls, echo=TRUE}
sol.stk <- readFLSss3(dir, name="SOL")
plot(sol.stk)
```

```{r solreadflsr, echo=TRUE}
sol.sr <- readFLSRss3(dir)
sol.sr
```

```{r solreadflrp, echo=TRUE}
sol.rp <- readFLRPss3(dir)
sol.rp
```

```{r solreadkobe, echo=TRUE}
sol.kobe <- readKobess3(dir)
sol.kobe
```

```{r solreadres, echo=TRUE}
sol.res <- readRESss3(dir)
head(sol.res)
```

```{r solreadflib, echo=TRUE}
sol.idx <- readFLIBss3(dir)
sol.idx
```

```{r solreadom, echo=TRUE}
sol.om <- readFLomss3(dir)
sol.om
```

```{r solreadoms, echo=TRUE}
sol.oms <- readOMSss3(dir)
names(sol.oms)
```

```{r solreadcd, echo=TRUE}
sol.cd <- readCDss3(dir, name="sol274")
names(sol.cd)
```

`readFLoemss3`, `readFLomeOMss3`, `readSRIss3` and `readRESIDss3` build objects used to run an observation error model as part of an MSE loop, but currently only work with **SS** models of version 3.24 or earlier, so they cannot be demonstrated on the `sol` (**SS** 3.30) dataset.

## `build` functions on `sol`

The `build*` functions take the list already loaded by `readOutputss3()`, or stored in `data(sol)`, and construct the corresponding FLR object. As `sol` comes from a **SS** 3.30 model, the `*ss330` variants must be used.

| Function          | Output class          | Description                          |
|-------------------|------------------------|---------------------------------------|
| buildFLSss330     | *FLStock*              | Stock and catch time series           |
| buildFLSRss3      | *FLSR*                 | Stock-recruitment relationship        |
| buildFLRPss330    | *FLPar*                | Reference points                      |
| buildKobess330    | *FLQuants*             | B/BMSY and F/FMSY time series          |
| buildRESss330     | *data.table*           | Time series of stock summary results  |
| buildFLIBss330    | *FLIndices*            | Abundance indices (`FLIndexBiomass`)  |
| buildFLIss330     | *FLIndices*            | Abundance indices (`FLIndex`)         |
| buildFLBss330     | *FLBiol*               | Biological time series                |
| buildFLBFss330    | *FLBiol* + *FLFisheries*| Biology and fisheries                |
| buildLCss330      | *FLQuant*              | Length composition data               |
| buildKobess3, buildFLRPss3, buildRESss3, buildFLIBss3, buildFLSss3 | | Equivalent builders for **SS** 3.24 models |

```{r solbuildfls, echo=TRUE}
sol.stk <- buildFLSss330(sol)
plot(sol.stk)
```

```{r solbuildflsr, echo=TRUE}
sol.sr <- buildFLSRss3(sol)
sol.sr
```

```{r solbuildflrp, echo=TRUE}
sol.rp <- buildFLRPss330(sol)
sol.rp
```

```{r solbuildkobe, echo=TRUE}
sol.kobe <- buildKobess330(sol)
sol.kobe
```

```{r solbuildres, echo=TRUE}
sol.res <- buildRESss330(sol)
head(sol.res)
```

```{r solbuildflib, echo=TRUE}
sol.idx <- buildFLIBss330(sol)
sol.idx
```

```{r solbuildfli, echo=TRUE}
sol.fli <- buildFLIss330(sol)
sol.fli
```

```{r solbuildflb, echo=TRUE}
sol.bio <- buildFLBss330(sol)
sol.bio
```

`buildFLBFss330`, which constructs an `FLBiol` plus a matching set of `FLFisheries`, and `buildLCss330`, which extracts length composition data from a `dat` input file, are not demonstrated here as they require, respectively, a fleet structure and length composition data not present in the `sol` model.

# Loading a full OM grid using `load`

# Flow of information SS3 - FLR

# Tested SS models

# Checking an operating model

- Convergence
- Estimates of B0
- Recruitment residual trends

# Auxiliary functions

- runtime
- convergencelevel,
- extractFbar, extractSSB, extractRec, extractZatage
- packss3run

